Analysis of variability

### 2016-08-13 portfolio/analysis/1-create-modelling-data.Rmd
Tidy file names / archive many files

cfecAnnual = group_by(cfecnew, p_holder, year) %>%
  summarize(permit = p_fshy[1], boat = cadfg[1],
    days = diff(range(yday) + 1),
    days_n = length(unique(yday)),
    days_permit = tot.permitDays[1], 
    ngear = length(unique(gearn)),
    nspec = length(unique(specn)),
    npermit = length(unique(p_fshy)),
    revenue = sum(g_earn),
    weight = sum(g_pounds),
    length = mean(vlength, na.rm=TRUE),
    strategy = strategy[1],
    strategy_gear = strategy_gear[1],
    strategy_permit = strategy.permit[1],
    chnk = sum(g_earn[spec=="CHNK"]),
    coho = sum(g_earn[spec=="COHO"]),
    pink = sum(g_earn[spec=="PINK"]),
    sock = sum(g_earn[spec=="SOCK"]),
    chum = sum(g_earn[spec=="CHUM"]))

@@

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
cfecAnnual$strategy_gear.prev = cfecAnnual$strategy_gear[idx]
cfecAnnual$strategy_permit.prev = cfecAnnual$strategy_permit[idx]
cfecAnnual$weight.prev = cfecAnnual$weight[idx]
cfecAnnual$days.prev = cfecAnnual$days[idx]
cfecAnnual$days_permit.prev = cfecAnnual$days_permit[idx]
cfecAnnual$chnk.prev = cfecAnnual$chnk[idx]
cfecAnnual$coho.prev = cfecAnnual$coho[idx]
cfecAnnual$chum.prev = cfecAnnual$chum[idx]
cfecAnnual$pink.prev = cfecAnnual$pink[idx]
cfecAnnual$sock.prev = cfecAnnual$sock[idx]
```

>>>>

cfecAnnual = group_by(cfecnew, p_holder, year) %>%
  summarize(permit = p_fshy[1], boat = cadfg[1],
    days = diff(range(yday) + 1),
    days_n = length(unique(yday)),
    days_permit = tot.permitDays[1], 
    ngear = length(unique(gearn)),
    nspec = length(unique(specn)),
    npermit = length(unique(p_fshy)),
    revenue = sum(g_earn),
    weight = sum(g_pounds),
    length = mean(vlength, na.rm=TRUE),
    strategy = strategy[1],
    strategy_gear = strategy_gear[1],
    strategy_permit = strategy.permit[1],
    chnk = sum(g_earn[spec=="CHNK"]),
    coho = sum(g_earn[spec=="COHO"]),
    pink = sum(g_earn[spec=="PINK"]),
    sock = sum(g_earn[spec=="SOCK"]),
    chum = sum(g_earn[spec=="CHUM"]),
    region = region[1],
    nregion = length(unique(region)))

@@

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
cfecAnnual$strategy_gear.prev = cfecAnnual$strategy_gear[idx]
cfecAnnual$strategy_permit.prev = cfecAnnual$strategy_permit[idx]
cfecAnnual$weight.prev = cfecAnnual$weight[idx]
cfecAnnual$days.prev = cfecAnnual$days[idx]
cfecAnnual$days_permit.prev = cfecAnnual$days_permit[idx]
cfecAnnual$chnk.prev = cfecAnnual$chnk[idx]
cfecAnnual$coho.prev = cfecAnnual$coho[idx]
cfecAnnual$chum.prev = cfecAnnual$chum[idx]
cfecAnnual$pink.prev = cfecAnnual$pink[idx]
cfecAnnual$sock.prev = cfecAnnual$sock[idx]

cfecAnnual$region.prev = cfecAnnual$region[idx]
cfecAnnual$nregion.prev = cfecAnnual$nregion[idx]
```

### 2016-08-13 portfolio/analysis/archive/7-separate-models.Rmd
Tidy file names / archive many files

```{r filter-data}
dm <- group_by(dm, p_holder) %>%
  mutate(nyr = length(unique(year))) %>%
  mutate(range_div = diff(range(specDiv))) %>%
  as_data_frame() %>%
  filter(nyr >= 5) %>%
  filter(range_div > 0)
nrow(dm)

@@

```{r condense-data-by-permit-holder}
dmv <- dm %>% group_by(p_holder) %>%
  summarise(cv = sd(revenue)/mean(revenue), log_specDiv = mean(log(specDiv)),
    # summarise(cv = mean(revenue), log_specDiv = mean(log(specDiv)),
    length = mean(log(length+1), na.rm = TRUE), days = mean(days, na.rm = TRUE),
    strategy = paste(sort(unique(substr(strategy, 1, 1))), collapse = " ")) %>%
  as_data_frame() %>%
  group_by(strategy) %>%
  mutate(n_strategy = n()) %>%
  filter(n_strategy > 4) %>%
  mutate(mean_log_spec = mean(log_specDiv)) %>%
  as_data_frame() %>%
  mutate(scaled_log_days = scale(log(days + 1)), scaled_log_spec = scale(log_specDiv),
    scaled_log_length = scale(length))
dmv$strategy <- as.factor(dmv$strategy)
```

@@

@@

>>>>

```{r filter-data}
dm <- group_by(dm, p_holder) %>%
  mutate(nyr = length(unique(year))) %>%
  mutate(range_div = diff(range(specDiv))) %>%
  as_data_frame() %>%
  filter(nyr >= 10) %>%
  filter(range_div > 0)
nrow(dm)

@@

```{r condense-data-by-permit-holder}
dmv <- dm %>% group_by(p_holder) %>%
    summarise(cv = sd(revenue)/mean(revenue), log_specDiv = mean(log(specDiv)),
    # summarise(cv = sd(log(revenue)-log(revenue.prev), na.rm = T),
      log_specDiv = mean(log(specDiv)),
    # summarise(cv = mean(revenue), log_specDiv = mean(log(specDiv)),
    length = mean(log(length+1), na.rm = TRUE), days = mean(days, na.rm = TRUE),
    strategy = paste(sort(unique(substr(strategy, 1, 1))), collapse = " ")) %>%
  as_data_frame() %>%
  filter(!is.na(cv)) %>%
  group_by(strategy) %>%
  mutate(n_strategy = n()) %>%
  filter(n_strategy > 4) %>%
  mutate(mean_log_spec = mean(log_specDiv)) %>%
  as_data_frame() %>%
  mutate(scaled_log_days = scale(log(days + 1)), scaled_log_spec = scale(log_specDiv),
    scaled_log_length = scale(length))
dmv$strategy <- as.factor(dmv$strategy)
```

@@

```{r}
library(mgcv)
dmv <- dmv %>% group_by(strategy) %>% mutate(cv_centered = log(cv) - mean(log(cv)))
m1 <- gam(cv_centered ~ s(scaled_log_spec) + s(scaled_log_days), data = dmv)
par(mfrow = c(1,2))
plot(m1)
```

@@

With differenced data

```{r}
df = readRDS(file="../data-generated/cfec-diff-for-modeling.rds")
dmv <- df %>% group_by(p_holder) %>%
    summarise(sd_ = sd(log(revenue/revenue.prev)), log_specDiv = mean(log(specDiv)),
    length = mean(log(length+1), na.rm = TRUE), days = mean(log(days+1), na.rm = TRUE),
    strategy = paste(sort(unique(substr(strategy, 1, 1))), collapse = " ")) %>%
  as_data_frame() %>%
  filter(!is.na(sd_)) %>%
  group_by(strategy) %>%
  mutate(n_strategy = n()) %>%
  filter(n_strategy > 6) %>%
  mutate(mean_log_spec = mean(log_specDiv)) %>%
  as_data_frame() %>%
  mutate(scaled_log_days = scale(days), scaled_log_spec = scale(log_specDiv),
    scaled_log_length = scale(length))
dmv$strategy <- as.factor(dmv$strategy)

library(mgcv)
dmv <- dmv %>% group_by(strategy) %>% mutate(log_sd = log(sd_) - mean(log(sd_)))
m1 <- gam(log_sd ~ s(scaled_log_spec) + s(scaled_log_days), data = dmv)
par(mfrow = c(1,2))
plot(m1)
